Signed Peripherality Functions for Multivariate Statistical Inference

### 2018-03-23 MS-new/3_Scale.tex
main paper v1 done, proofs and intro remain

\section{Inverse depth and estimation of scatter}
\label{section:dcmSection}
We elaborate on the problem estimating the covariance matrix $\Sigma$ and its eigenvectors and eigenvalues using signed peripherality functions in this section. In addition to conditions (P) and (A1), we assume the following conditions on these functions in this section:

\vspace{1em}
\noindent\textbf{(B1)} The following holds:
$$
P(\bfx; \BF) \leq P(\bfmu_F + t(\bfx - \bfmu_F), \BF),
$$
%
for any $\bfx \in \BR^p$ and $t \in (0,1)$.

@@

Given an inverse depth function $D^-_{\bfX}(\bfx)$ we transform the original random variable as: $\tilde \bfX = D^-_\bfX(\bfx) \bfS(\bfx - \bfmu)$. We attempt to use these vector as the proxy of original data for improved estimation of the components of the population covariance matrix $\Sigma$ in this section. To this end, in Section~\ref{subsec:dcm} we give some results characterizing the population covariance matrix of these vectors, i.e. $\BV \tilde \bfX$, and its affine equivariant counterpart.

@@

\subsection{Depth Covariance Matrix}
\label{subsec:dcm}
Consider the spectral decomposition of $\Sigma$: $\Sigma = \Gamma\Lambda\Gamma^T$, $\Gamma$ being orthogonal and $\Lambda$ diagonal with positive diagonal elements $\lambda_1, \ldots, \lambda_p$. Also normalize the original random variable as $\bfZ = \Gamma^T\Lambda^{-1/2} (\bfX - \bfmu)$. In this setup, we can represent the transformed random variable as
%
\begin{eqnarray}
\tilde \bfX &=& D^-_{\bfX} (\bfX) \bfS(\bfx - \bfmu) \notag \\
&=& D^-_{\Gamma\Lambda^{1/2}\bfZ + \bfmu} (\Gamma\Lambda^{1/2} \bfZ + \bfmu). \bfS(\Gamma\Lambda^{1/2} \bfZ) \notag \\
&=& \Gamma D^-_{\bfZ}(\bfZ) \bfS(\Lambda^{1/2}\bfZ) \notag \\
&=& \Gamma \Lambda^{1/2} D^-_{\bfZ}(\bfZ) \bfS(\bfZ) \frac{\| \bfZ \|}{\|\Lambda^{1/2} \bfZ \|}.
\label{equation:rankdecomp}
\end{eqnarray}
%
%Because of affine (thus rotational) invariance of a depth function, the depth (htped) value at $\bfz$ does not depend on the direction of $\bfz$, i.e. $D^-_{\bfZ}(\bfz)$ and $\bfS(\bfz)$ are independent. Furthermore,
%$$ Cov \left(\bfS (\bfz), \frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) = E \left(\bfS (\bfz). \frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) - E \bfS (\bfz) E \left(\frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) = E \left( \frac{\bfz}{\|\Lambda^{1/2} \bfz\|} \right) = \bf0$$
%both $\bfS (\bfz)$ and $\bfz / \| \Lambda^{1/2}\bfz \|$ are odd functions of $\bfz$, which has a circularly symmetric distribution, hence each of them has expectation $\bf0$. Consequently, we obtain an expression for the covariance matrix of $\tilde \bfX$:

$D^-_\bfZ(\cdot)$ is an even function in its argument because of affine invariance, as is $\| \bfz \| / \|\Lambda^{1/2} \bfz \|$. Since the sign function $\bfS(\cdot)$ is odd in its argument, it follows that $\BE \tilde \bfX = \bf0$. Consequently, we obtain an expression for $\BV \tilde \bfX$, which we call the {\it Depth Covariance Matrix} (DCM):

\begin{Theorem} \label{Theorem:covform}
Let the random variable $\bfX \in \mathbb{R}^p$ follow an elliptical distribution with center $\bfmu$ and covariance matrix $\Sigma = \Gamma\Lambda\Gamma^T$, its spectral decomposition. Then, given a depth function $D_\bfX(.)$ the covariance matrix of the transformed random variable $\tilde\bfX$ is
\begin{equation} \label{equation:covformEq1}
\BV \tilde \bfX) = \Gamma \Lambda_{D} \Gamma^T,\quad\mbox{with}\quad \Lambda_{D} = \mathbb E_\bfZ \left[ (D^-_\bfZ(\bfz))^2 \frac{\Lambda^{1/2} \bfz \bfz^T \Lambda^{1/2}}{\bfz^T \Lambda \bfz} \right],
\end{equation}
where $\bfZ = (Z_1,...,Z_p)^T \sim N({\bf 0}, \bfI_p)$, so that $\Lambda_{D}$ a diagonal matrix with diagonal entries
%
$$
\lambda_{D,i} = \mathbb E_\bfZ \left[ \frac{(D^-_\bfZ(\bfz))^2 \lambda_i z_i^2}{\sum_{j=1}^p \lambda_j z_j^2} \right].
$$
\end{Theorem}

The matrix of eigenvectors $\Gamma$ remains unchanged in the transformation $\bfX \rightarrow \tilde \bfX$. As a result, the multivariate rank vectors can be used for robust principal component analysis, which we are going to discuss shortly. However, as one can see in the above expression, the diagonal entries of $\Lambda_{D}$ do not change if a scale change is done on all entries of $\Lambda$, meaning the $\Lambda_{D}$ matrices corresponding to $\BF$ and $c \BF$ for some $c \neq 0$ will be same. This is the reason the DCM is not equivariant under affine transformations.

We need to follow the general framework of M-estimation with data-dependent weights \citep{HuberBook81} to construct an affine equivariant counterpart of the DCM. Specifically, we implicitly define the Affine-equivariant Depth Covariance Matrix (ADCM) as
%
\begin{equation} \label{eqn:ADCM}
\Sigma_{Dw} = \frac{1}{ \BV \tilde Z_1 } \BE_\bfX \left[ \frac{(D^-_\bfX(\bfx))^2 (\bfx - \bfmu) (\bfx - \bfmu)^T}{(\bfx - \bfmu)^T \Sigma_{Dw}^{-1} (\bfx - \bfmu)} \right].
\end{equation}
%
Its affine equivariance follows from the fact that the weights $(D^-_\bfX (\bfx))^2$ depend only on the standardized quantities $\bfz \sim \cE ({\bf 0}, k \bfI_p, g)$. We solve (\ref{eqn:ADCM}) iteratively by obtaining a sequence of positive definite matrices $\Sigma^{(k)}_{Dw}$ until convergence:
%
$$
\Sigma^{(k+1)}_{Dw} = \frac{1}{\BV \tilde Z_1 } \BE_\bfX \left[ \frac{(D^-_\bfX(\bfx))^2 (\Sigma^{(k)}_{Dw})^{1/2} (\bfx - \bfmu) (\bfx - \bfmu)^T (\Sigma^{(k)}_{Dw})^{1/2}}{(\bfx - \bfmu)^T (\Sigma^{(k)}_{Dw})^{-1} (\bfx - \bfmu)} \right].
$$
%

To ensure existence and uniqueness of this estimator, let us consider the class of scatter estimators $\Sigma_M$ that are obtained as solutions of the following equation:
%
\begin{equation}
E_{\bfZ_M} \left[ u( \| \bfz_M \| )  \frac{\bfz_M \bfz_M^T}{\| \bfz_M \|^2}  - v( \| \bfz_M \| ) I_p \right] = 0
\end{equation}
%
with $\bfz_M = \Sigma_M^{-1/2} (\bfx - \bfmu)$. Under the following assumptions on the scalar valued functions $u$ and $v$, the above equation produces a unique solution \citep{HuberBook81}:
%

\vspace{1em}
\noindent\textbf{(M1)} The function $u(r)/r^2$ is monotone decreasing, and $u(r)>0$ for $r>0$;

\noindent\textbf{(M2)}  The function $v(r)$ is monotone decreasing, and $v(r)>0$ for $r>0$;

\noindent\textbf{(M3)} Both $u(r)$ and $v(r)$ are bounded and continuous;

\noindent\textbf{(M4)} $u(0) / v(0) < p$;

\noindent\textbf{(M5)} For any hyperplane in the sample space $\mathcal X$, (i) $P(H) = E_\bfX 1_{\bfx \in H} < 1 - p v(\infty) / u(\infty)$ and (ii) $P(H) \leq 1/p$.
%

\vspace{1em}
\noindent In our case we take $v(r) = Var(\tilde Z_1)$, i.e. a constant, thus (M2) and (M3) are trivially satisfied. As for $u$, we notice that most well-known depth functions can be expressed as simple functions of the norm of the standardized random variable. For example, $PD_\bfZ (\bfz) = (1 - G(\| \bfz \|); MhD_\bfZ (\bfz) = (1+\| \bfz \|^2)^{-1}; HSD_\bfZ (\bfz) = (1+\| \bfz \|)^{-1}$ etc., so that we can take as $u$ square of the corresponding peripherality functions:
% Var(Z1) breaks into two independent parts: depth and sign. So can check m4 and M5 here itself.
$$
u_{PD} (r) = G^2 (r); \quad u_{MhD}(r)  = \frac{r^4}{(1 + r^2)^2}; \quad u_{HSD}(r)  = \frac{r^2}{(1 + r/G^{-1}(0.75))^2}
$$
%

It is easy to verify that the above choices of $u$ satisfy (M1) and (M3). To check (M4) and (M5), first notice that $\bfZ$ has a spherically symmetric distribution, so that its norm and sign are independent. Since $D_\bfZ(\bfz)$ depends only on $\| \bfz \|$, we have
%
$$
Var (\tilde Z_1) = Var \left( D^-_\bfZ (\bfZ) \frac{Z_1}{ \|\bfZ \|} \right) = Var (D^-_\bfZ (\bfZ)) Var (S_1 (\bfZ)) = \frac{1}{p} Var (D^-_\bfZ (\bfZ))
$$
%
as $Cov(\bfS(\bfZ)) = Cov((S_1(\bfZ), S_2(\bfZ), ..., S_p(\bfZ))^T) = I_p/p$. Now for MhD and HSD $u(\infty)=1, u(0)=0$, so (M4) and (M5) are immediate. To achieve this for PD, we only need to replace $u_{PD}(r)$ with $u_{PD}^*(r) = G^2(r) - 1/4$.

\subsection{Calculating the sample DCM and ADCM}
Let us now consider $n$ iid random draws from our elliptic distribution $F$, say $\bfX_1,...,\bfX_n$. For ease of notation, denote $SS(\bfx; \bfmu) = \bfS(\bfx - \bfmu) \bfS(\bfx - \bfmu)^T$. Then, given the depth function and known location center $\bfmu$, one can show that the vectorized form of $\sqrt n$-times the sample DCM: $\sum_{i=1}^n (D^-_\bfX(\bfx_i))^2SS(\bfx_i; \bfmu) /\sqrt{n}$  has an asymptotic multivariate normal distribution with mean $\sqrt n.vec(E[( (D^-_\bfX (\bfX))^2 SS(\bfx; \bfmu)])$ and a certain covariance matrix by straightforward application of the central limit theorem (CLT). But in practice the population depth function $D_\bfX(\bfx) = D(\bfx, F)$ is estimated by the depth function based on the empirical distribution function, $F_n$. Denote this sample depth by $D^n_\bfX (\bfx) = D(\bfx, F_n)$. Here we make the following assumption regarding how it approximates $D_\bfX(\bfx)$:

\vspace{1em}
\noindent\textbf{(D5)} \textit{Uniform convergence}: $\sup_{\bfx \in \mathbb R^p} | D^n_\bfX (\bfx) - D_\bfX (\bfx) | \rightarrow 0$ as $n \rightarrow \infty $.
\vspace{1em}

The assumption that empirical depths converge uniformly at all points $\bfx$ to their population versions holds under very mild conditions for several well known depth functions: for example projection depth \citep{zuo03} and simplicial depth \citep{Dumbgen92}. One also needs to replace the known location parameter $\bfmu$ by some estimator $\hat\bfmu_n$. Examples of robust estimators of location that are relevant here include the spatial median \citep{haldane48,brown83}, Oja median \citep{oja83}, projection median \citep{zuo03} etc. Now, given $D^n_\bfX(.)$ and $\hat \bfmu_n$, to plug them into the sample DCM and still go through with the CLT we need the following result:

\begin{Lemma} \label{Lemma:lemma1}
Consider a random variable $\bfX \in \mathbb{R}^p$ having a continuous and symmetric distribution with location center $\bfmu$ such that $E\|\bfx - \bfmu \|^{-3/2} < \infty$. Given $n$ random samples from this distribution, suppose $\hat\bfmu_n$ is an estimator of $\bfmu$ so that $\sqrt n (\hat\bfmu_n - \bfmu) = O_P(1) $. Then with the above notations, and given the assumption (D5) we have
%
$$ \sqrt n \left[
\frac{1}{n} \sum_{i=1}^n (D^{-n}_\bfX (\bfx_i))^2 SS(\bfx_i; \hat\bfmu_n) -
\frac{1}{n} \sum_{i=1}^n (D^-_\bfX (\bfx_i))^2 SS(\bfx_i; \bfmu) \right]
\stackrel{P}{\rightarrow} 0 $$
\end{Lemma}

Following this, we are now in a position to state the result for consistency of the sample DCM:

\begin{Theorem} \label{Theorem:rootn}
Consider $n$ iid samples from the distribution in Lemma \ref{Lemma:lemma1}. Then, given a depth function $D_\bfX(.)$ and an estimate of center $\hat\bfmu_n$ so that $\sqrt n(\hat \bfmu_n - \bfmu) = O_P(1)$,
%
$$ \sqrt n \left[ vec\left\{ \frac{1}{n} \sum_{i=1}^n (D^{-n}_\bfX (\bfx_i))^2 SS(\bfx_i; \hat\bfmu_n) \right\} - E \left[ vec\left\{ (D^-_\bfX (\bfx))^2 SS(\bfx; \bfmu) \right\} \right] \right]
\stackrel{D}{\rightarrow}
N_{p^2} ({\bf 0}, V_{D,S}(F)) $$
$$ \text{with } V_{D,S}(F) = Var \left[vec \left\{ (D^-_\bfX (\bfx))^2 SS(\bfx; \bfmu) \right\} \right] $$
\end{Theorem}

\noindent In case $F$ is elliptical, an elaborate form of the covariance matrix $V_{D,S}(F)$ explicitly specifying each of its elements (more directly those of its $\Gamma^T$-rotated version) can be obtained, which is given in Appendix \ref{section:appA}. This form is useful when deriving limiting distributions of eigenvectors and eigenvalues of the sample DCM.

In contrast to the DCM, the issue of estimating $\bfmu$ to plug into the ADCM is easily handled by simultaneously solving for the location and scatter functionals ($\bfmu_{Dw}, \Sigma_{Dw}$):
%
\begin{eqnarray}
E \left[ \frac{\Sigma_{Dw}^{-1/2} (\bfx - \bfmu_{Dw})}{ \| \Sigma_{Dw}^{-1/2}(\bfx - \bfmu_{Dw}) \|} \right] &=& {\bf 0}_p \label{eqn:AffineMedian}\\
E \left[ \frac{(D^-_\bfX(\bfx))^2 \Sigma_{Dw}^{-1/2} (\bfx - \bfmu_{Dw}) (\bfx - \bfmu_{Dw})^T \Sigma_{Dw}^{-1/2}}{(\bfx - \bfmu_{Dw})^T \Sigma_{Dw}^{-1} (\bfx - \bfmu_{Dw})} \right] &=& Var (\tilde Z_1) I_p \label{eqn:AffineCov}
\end{eqnarray}
%
In the framework of (\ref{eqn:ADCM}), for any fixed $\Sigma_M$ there exists a unique and fixed solution of the location problem $E_{\bfZ_M} (w(\| \bfz_M \| \bfz_M) = {\bf 0}_p $ under the following condition:

\vspace{1em}
\noindent\textbf{(M6)} The function $w(r)r$ is monotone increasing for $r>0$.

\vspace{1em}
\noindent This condition is easy to verify for our choice of the weights: $w(\| \bfz_M \| ) = D^-_{\bfZ_M} (\bfz_M) / \| \bfz_M \|$. Uniqueness of simultaneous fixed point solutions of \ref{eqn:AffineMedian} and \ref{eqn:AffineCov} is guaranteed when $\bfX$ has a symmetric distribution \citep{HuberBook81}.

In practice it is difficult to calculate the scale multiple $Var (\tilde Z_1)$ analytically for known depth functions and an arbitrary $F$. Here we instead obtain its standardized version $\Sigma_{Dw}^* = \Sigma_{Dw} / Var(\tilde Z_1)$ (so that the determinant equals 1), alongwith $\bfmu_{Dw}$ using the following iterative algorithm:

\begin{enumerate}
\item Start from some initial estimates $(\bfmu_{Dw}^{(0)}, \Sigma_{Dw,(0)})$. Set $t=0$;

\item Calculate the standardized observations $\bfz_i^{(t)} = \Sigma_{Dw,(t)}^{-1/2} (\bfx_i - \bfmu_{Dw}^{(t)})$;

\item Update the location estimate:
%
$$
\bfmu_{Dw}^{(t+1)} = \frac{\sum_{i=1}^n \tilde \bfx_i / \| \bfz_i^{(t)} \| }{\sum_{i=1}^n 1 / \| \bfz_i^{(t)} \|}
$$
%
\item Update the scatter estimate:
%
$$
\Sigma_{Dw}^{*(t+1)} = \frac{1}{n} \sum_{i=1}^n \frac{(D^{-n}_\bfX (\bfx_i))^2 (\bfx_i - \bfmu_{Dw}^{(t+1)})(\bfx_i - \bfmu_{Dw}^{(t+1)})^T}{\| \bfz_i^{(t)} \|^2}; \quad \Sigma_{Dw}^{*(t+1)} \leftarrow \frac{\Sigma_{Dw}^{*(t+1)}}{\text{det} (\Sigma_{Dw}^{*(t+1)})^{1/p}}
$$
%
\item Continue until convergence.
\end{enumerate}

>>>>

\section{Inverse depth and estimation of scatter}
\label{section:dcmSection}
We focus on the estimation of scatter functionals of $\BF$ using peripherality-weighted sign vectors in this section. To this end, in addition to conditions (P) and (A1), we assume the following conditions on the peripherality functions:

\vspace{1em}
\noindent\textbf{(B1)} The following holds for any $\bfx \in \BR^p$ and $t \in (0,1)$:
%
$$
P(\bfx; \BF) \leq P(\bfmu_F + t(\bfx - \bfmu_F), \BF),
$$
%

@@

Given an inverse depth function $D^-_{\bfX}(\bfx)$ we transform the original random variable as: $\tilde \bfX = D^-_\bfX(\bfx) \bfS(\bfx - \bfmu)$. We attempt to use the transformed data matrix $\tilde \BX = (\tilde \bfx_1, \ldots, \tilde \bfx_n)^T$ as proxy of the original data matrix $\BX$ for improved estimation of the components of the population covariance matrix $\Sigma$ in this section. To this end, in Section~\ref{subsec:dcm} we give some results characterizing the population covariance matrix of these vectors, i.e. $\BV \tilde \bfX$, and its affine equivariant counterpart. {\colrbf details of 3.2 and rest}

@@

\subsection{Depth Covariance Matrix}
\label{subsec:dcm}
Consider the spectral decomposition of $\Sigma$: $\Sigma = \Gamma\Lambda\Gamma^T$, $\Gamma$ being orthogonal and $\Lambda$ diagonal with positive diagonal elements $\lambda_1 \geq \ldots \geq \lambda_p$. Also normalize the original random variable as $\bfZ = \Gamma^T\Lambda^{-1/2} (\bfX - \bfmu)$. In this setup, we can represent the transformed random variable as
%
\begin{eqnarray}
\tilde \bfX &=& D^-_{\bfX} (\bfX) \bfS(\bfX - \bfmu) \notag \\
&=& D^-_{\Gamma\Lambda^{1/2}\bfZ + \bfmu} (\Gamma\Lambda^{1/2} \bfZ + \bfmu). \bfS(\Gamma\Lambda^{1/2} \bfZ) \notag \\
&=& \Gamma D^-_{\bfZ}(\bfZ) \bfS(\Lambda^{1/2}\bfZ) \notag \\
&=& \Gamma \Lambda^{1/2} D^-_{\bfZ}(\bfZ) \bfS(\bfZ) \frac{\| \bfZ \|}{\|\Lambda^{1/2} \bfZ \|}.
\label{equation:rankdecomp}
\end{eqnarray}
%
%Because of affine (thus rotational) invariance of a depth function, the depth (htped) value at $\bfz$ does not depend on the direction of $\bfz$, i.e. $D^-_{\bfZ}(\bfz)$ and $\bfS(\bfz)$ are independent. Furthermore,
%$$ Cov \left(\bfS (\bfz), \frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) = E \left(\bfS (\bfz). \frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) - E \bfS (\bfz) E \left(\frac{\| \bfz \|}{\|\Lambda^{1/2} \bfz\|} \right) = E \left( \frac{\bfz}{\|\Lambda^{1/2} \bfz\|} \right) = \bf0$$
%both $\bfS (\bfz)$ and $\bfz / \| \Lambda^{1/2}\bfz \|$ are odd functions of $\bfz$, which has a circularly symmetric distribution, hence each of them has expectation $\bf0$. Consequently, we obtain an expression for the covariance matrix of $\tilde \bfX$:

$D^-_\bfZ(\cdot)$ is an even function in its argument because of affine invariance, as is $\| \bfz \| / \|\Lambda^{1/2} \bfz \|$. Since the sign function $\bfS(\cdot)$ is odd in its argument, it follows that $\BE \tilde \bfX = \bf0$. Consequently, we obtain an expression for $\tilde \Sigma := \BV \tilde \bfX$, which we call the {\it Depth Covariance Matrix} (DCM).

\begin{Theorem} \label{Theorem:covform}
Let the random variable $\bfX \in \mathbb{R}^p$ follow an elliptical distribution with center $\bfmu$ and covariance matrix $\Sigma = \Gamma\Lambda\Gamma^T$, its spectral decomposition. Then, given a depth function $D_\bfX(.)$ the covariance matrix of the transformed random variable $\tilde\bfX$ is
\begin{equation} \label{equation:covformEq1}
\tilde \Sigma = \Gamma \Lambda_{D} \Gamma^T,\quad\mbox{with}\quad \Lambda_{D} = \mathbb E_\bfZ \left[ (D^-_\bfZ(\bfz))^2 \frac{\Lambda^{1/2} \bfz \bfz^T \Lambda^{1/2}}{\bfz^T \Lambda \bfz} \right],
\end{equation}
where $\bfZ = (Z_1,...,Z_p)^T \sim N({\bf 0}, \BI_p)$, so that $\Lambda_{D}$ a diagonal matrix with diagonal entries
%
$$
\lambda_{D,i} = \mathbb E_\bfZ \left[ \frac{(D^-_\bfZ(\bfz))^2 \lambda_i z_i^2}{\sum_{j=1}^p \lambda_j z_j^2} \right].
$$
\end{Theorem}

The matrix of eigenvectors $\Gamma$ remains unchanged in the transformation $\bfX \rightarrow \tilde \bfX$. Thus the multivariate rank vectors can be used for robust principal component analysis (which we discuss shortly). However, as seen in \eqref{equation:covformEq1}, the diagonal entries of $\Lambda_{D}$ do not change if a scale change is done on all entries of $\Lambda$, meaning the $\Lambda_{D}$ matrices corresponding to $\BF$ and $c \BF$ for some $c \neq 0$ are same. Because of this reason, the DCM is not equivariant under affine transformations.

We follow the general framework of M-estimation with data-dependent weights \citep{HuberBook81} to construct an affine equivariant counterpart of the DCM. Specifically, we implicitly define the Affine-equivariant Depth Covariance Matrix (ADCM) as
%
\begin{equation} \label{eqn:ADCM}
\Sigma_{Dw} = \frac{1}{ \BV \tilde Z_1 } \BE_\bfX \left[ \frac{(D^-_\bfX(\bfx))^2 (\bfx - \bfmu) (\bfx - \bfmu)^T}{(\bfx - \bfmu)^T \Sigma_{Dw}^{-1} (\bfx - \bfmu)} \right].
\end{equation}
%
Its affine equivariance follows from the fact that the weights $(D^-_\bfX (\bfx))^2$ depend only on the standardized quantities $\bfz \sim \cE ({\bf 0}, k \bfI_p, g)$. We solve (\ref{eqn:ADCM}) iteratively by obtaining a sequence of positive definite matrices $\Sigma^{(k)}_{Dw}$ until convergence:
%
$$
\Sigma^{(k+1)}_{Dw} = \frac{1}{\BV \tilde Z_1 } \BE_\bfX \left[ \frac{(D^-_\bfX(\bfx))^2 (\Sigma^{(k)}_{Dw})^{1/2} (\bfx - \bfmu) (\bfx - \bfmu)^T (\Sigma^{(k)}_{Da})^{1/2}}{(\bfx - \bfmu)^T (\Sigma^{(k)}_{Dw})^{-1} (\bfx - \bfmu)} \right].
$$
%

To ensure existence and uniqueness of this estimator, we consider the class of scatter estimators $\Sigma_M$ that are obtained as solutions of the following equation:
%
\begin{equation}
E_{\bfZ_M} \left[ u( \| \bfz_M \| )  \frac{\bfz_M \bfz_M^T}{\| \bfz_M \|^2}  - v( \| \bfz_M \| ) \BI_p \right] = 0
\end{equation}
%
with $\bfz_M = \Sigma_M^{-1/2} (\bfx - \bfmu)$. Under the following assumptions on the scalar valued functions $u$ and $v$, the above equation produces a unique solution \citep{HuberBook81}:
%

\vspace{1em}
\noindent\textbf{(C1)} The function $u(r)/r^2$ is monotone decreasing, and $u(r)>0$ for $r>0$;

\noindent\textbf{(C2)}  The function $v(r)$ is monotone decreasing, and $v(r)>0$ for $r>0$;

\noindent\textbf{(C3)} Both $u(r)$ and $v(r)$ are bounded and continuous;

\noindent\textbf{(C4)} $u(0) / v(0) < p$;

\noindent\textbf{(C5)} For any hyperplane in the sample space $\mathcal X$, (i) $P(H) = \BE_\bfX \{ \BI{\bfx \in H} \} < 1 - p v(\infty) / u(\infty)$ and (ii) $P(H) \leq 1/p$.
%

\vspace{1em}
\noindent Putting things into context, in our case we have $v(r) = \BV\tilde Z_1, u(r) = (D^-_\bfX(\bfx))^2 = (D^-_\bfZ(\bfz))^2$. Since $v(r)$ is a constant, (C2) and (C3) are trivially satisfied. As for $u$, most well-known depth functions can be expressed as simple functions of the norm of the standardized random variable. For example, for Halfspace Depth (HSD) we have $D_\bfZ (\bfz) = (1 - G(\| \bfz \|)$, for Mahalanobis Depth (MhD) we have $D_\bfZ (\bfz) = (1+\| \bfz \|^2)^{-1}$ and for Projection Depth (PD) we have $D_\bfZ (\bfz) = (1+\| \bfz \|//MAD(G))^{-1}$, where $G$ is the cumulative distribution function of $g$, and MAD is median absolute deviation. Combined with the choice of inverse transformation, we get
% Var(Z1) breaks into two independent parts: depth and sign. So can check m4 and M5 here itself.
$$
u_{HSD} (r) = G^2 (r); \quad u_{MhD}(r)  = \frac{r^4}{(1 + r^2)^2}; \quad u_{PD}(r)  = \frac{r^2}{(1 + r/MAD(G))^2}.
$$
%
It is easy to verify that the above choices of $u$ satisfy (C1) and (C3). To check (C4) and (C5), notice that $\bfZ$ has a spherically symmetric distribution, so that its euclidean norm and spatial sign are independent. Since $D_\bfZ(\bfz)$ depends only on $\| \bfz \|$, we now have
%
$$
\BV \tilde Z_1 = \BV \left( D^-_\bfZ (\bfZ) \frac{Z_1}{ \|\bfZ \|} \right) = \BV (D^-_\bfZ (\bfZ)) \BV (S_1 (\bfZ)) = \frac{1}{p} \BV (D^-_\bfZ (\bfZ)),
$$
%
as $\BV(\bfS(\bfZ)) = \BV((S_1(\bfZ), S_2(\bfZ), ..., S_p(\bfZ))^T) = \BI_p/p$. Now for MhD and PD $u(\infty)=1, u(0)=0$, so (C4) and (C5) are immediate. To achieve this for HSD, we only need to replace $u_{HSD}(r)$ with $u_{HSD*}(r) = G^2(r) - 1/4$.

\subsection{Calculating the sample DCM and ADCM}
Let us denote $\BS(\bfx; \bfmu) = \bfS(\bfx - \bfmu) \bfS(\bfx - \bfmu)^T$. Then, given the depth function and {\it known} location center $\bfmu$, one can show that the vectorized form of $\sqrt n$-times the sample DCM, i.e. $n^{-1/2} \sum_{i=1}^n (D^-_\bfX(\bfx_i))^2 \BS (\bfx_i; \bfmu)$  has an asymptotic multivariate normal distribution with mean $\sqrt n. \ve( \BE [( (D^-_\bfX (\bfX))^2 \BS(\bfx; \bfmu)])$ and some covariance matrix by straightforward application of the central limit theorem (CLT). But in practice the depth at any point $\bfx$ is estimated by the sample depth, i.e. depth calculated using empirical distribution function in place of $\BF$. We also need to replace the known location parameter $\bfmu$ by some estimator $\hat\bfmu_n$.

Denote this sample depth by $D^n_\bfX (\bfx)$, and the inverse depth by $D^{n-}_\bfX (\bfx)$. Here we make the following assumption regarding the sample depths:

\vspace{1em}
\noindent\textbf{(D1)} \textit{Uniform convergence}: $\sup_{\bfx \in \mathbb R^p} | D^n_\bfX (\bfx) - D_\bfX (\bfx) | \rightarrow 0$ as $n \rightarrow \infty $.
\vspace{1em}

\noindent The above holds under very mild conditions for several depth functions, e.g. projection depth \citep{zuo03} and simplicial depth \citep{Dumbgen92}. For $\hat \bfmu_n$ we use a robust estimate of location, examples of which include the spatial median \citep{haldane48,brown83}, Oja median \citep{oja83} and projection median \citep{zuo03}.

To plug in the above estimates to the sample DCM and go through with the CLT, we need the following result.

\begin{Lemma} \label{Lemma:lemma1}
Consider a random variable $\bfX \in \mathbb{R}^p$ having a continuous and symmetric distribution with location center $\bfmu$ such that $\BE \|\bfx - \bfmu \|^{-3/2} < \infty$. For a size-$n$ random sample from this distribution, suppose $\hat\bfmu_n$ is an estimator of $\bfmu$ so that $\hat\bfmu_n - \bfmu = O_P(n^{-1/2}) $. Then, given assumption (D1) we have
%
$$
\frac{1}{n} \sum_{i=1}^n (D^{n-}_\bfX (\bfx_i))^2 \BS(\bfx_i; \hat\bfmu_n) = \frac{1}{n} \sum_{i=1}^n (D^-_\bfX (\bfx_i))^2 \BS(\bfx_i; \bfmu)
+ \BR_n,
$$
%
where $\| \BR_n \|_\infty = o_P(n^{-1/2})$.
\end{Lemma}

\noindent We are now in a position to state the result for consistency of the sample DCM:

\begin{Theorem} \label{Theorem:rootn}
Consider a size-$n$ random sample from the distribution in Lemma \ref{Lemma:lemma1}. Then, given a depth function $D_\bfX(.)$ and an estimate of center $\hat\bfmu_n$ so that $\sqrt n(\hat \bfmu_n - \bfmu) = O_P(1)$,
%
$$ \ve \left\{ \frac{1}{n} \sum_{i=1}^n (D^{n-}_\bfX (\bfx_i))^2 \BS(\bfx_i; \hat\bfmu_n) \right\} -
\BE [ \ve \left\{ (D^-_\bfX (\bfx))^2 \BS(\bfx; \bfmu) \right\} ] \rightsquigarrow
\cN_{p^2} ({\bf 0}, \BV_D(\BF)) + \bfR_n,
$$
%
with $\BV_D(\BF) = \BV [\ve \{ (D^-_\bfX (\bfx))^2 \BS(\bfx; \bfmu) \} ] $ and $\| \bfR_n \|_\infty = o_P(n^{-1/2})$.
\end{Theorem}

\noindent When $\BF$ is elliptical, an elaborate form of the covariance matrix $\BV_D(\BF)$ explicitly specifying each of its elements (more directly those of its $\Gamma^T$-rotated version) can be obtained (see Appendix \ref{section:appA}). This form is useful when deriving limiting distributions of eigenvectors and eigenvalues of the sample DCM.

For ADCM, we simultaneously estimate its sample version and the mean vector $\bfmu$ by solving for the location and scatter functionals ($\bfmu_{Dw}, \Sigma_{Dw}$):
%
\begin{eqnarray}
\BE \left[ \frac{\Sigma_{Dw}^{-1/2} (\bfx - \bfmu_{Dw})}{ \| \Sigma_{Dw}^{-1/2}(\bfx - \bfmu_{Dw}) \|} \right] &=& {\bf 0}_p, \label{eqn:AffineMedian}\\
\frac{1}{\BV \tilde Z_1}
\BE \left[ \frac{(D^-_\bfX(\bfx))^2 \Sigma_{Dw}^{-1/2} (\bfx - \bfmu_{Dw}) (\bfx - \bfmu_{Dw})^T \Sigma_{Dw}^{-1/2}}{(\bfx - \bfmu_{Dw})^T \Sigma_{Dw}^{-1} (\bfx - \bfmu_{Dw})} \right] &=& \BI_p. \label{eqn:AffineCov}
\end{eqnarray}
%
In the M-estimation framework of \cite{HuberBook81}, for any fixed $\Sigma_M$ there exists a unique and fixed solution of the general location problem $\BE_{\bfZ_M} (w(\| \bfz_M \| \bfz_M) = {\bf 0}_p $ under the following condition:

\vspace{1em}
\noindent\textbf{(D2)} The function $w(r)r$ is monotone increasing for $r>0$.

\vspace{1em}
\noindent This is easy to verify for our choice of the weights: $w(\| \bfz_M \| ) = D^-_{\bfZ_M} (\bfz_M) / \| \bfz_M \|$. Uniqueness of simultaneous fixed point solutions of \eqref{eqn:AffineMedian} and \eqref{eqn:AffineCov} is guaranteed when $\bfX$ has a symmetric distribution \citep{HuberBook81}, which we also satisfy since $\bfX \sim \BF$, an elliptical distribution,

In practice, it is difficult to calculate the scale multiple $\BV \tilde Z_1$ analytically for known depth functions and an arbitrary $\BF$. We bypass this by instead obtaining a standardized version of the scatter functional: $\Sigma_{Dw}^* = \Sigma_{Dw} / \BV \tilde Z_1$ (and $\bfmu_{Dw}$) using the following iterative algorithm:

\begin{enumerate}
\item Start from some initial estimates $(\bfmu_{Dw}^{(0)}, \Sigma_{Dw}^{(0)}$. Set $t=0$;

\item Calculate the standardized observations $\bfz_i^{(t)} = (\Sigma_{Dw}^{(t)})^{-1/2} (\bfx_i - \bfmu_{Dw}^{(t)})$;
